Quantifying and controlling biases in dark matter halo concentration estimates

### 2016-08-25 paper/ms.tex
nailed the word count

We then use an Affine Invariant Markov Chain Monte Carlo (MCMC)
implemented in the python module {\texttt{emcee}} \citep{emcee} to sample the
likelihood function distribution.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty. 
We stress that different choices for $\sigma^2_{\rm eff}$ do not
affect the optimal concentration value, only its uncertainty.

@@

Figure \ref{fig:downsampling} summarizes our results.  
The plot on the left shows the average value of $f_{\rm off}$ as a
function of particle number.  
This can be interpreted as the statistical bias on the concentration
estimate. 
For large enough particle numbers, $N_{s}>4\times 10^3$ the results of
the three algorithms show a bias below the $1\%$ level.
For a lower number of particles the results start to deviate.
At $200$ particles the velocity method overestimates the concentration
by a factor $14\%$ while the density method overestimates it by $20\%$.
Around the same sampling scale, the new algorithm shows a more stable
behaviour underestimating the concentration only by a factor of
$1\%$-$3\%$.
Work  that studied the effect of
lowering numerical resolution on the same individual halos
\citep{2008MNRAS.391.1685S} yielded similar values for the
concentration across resolutions.  That work  explored a particle
number range $>5\times 10^5$, completely consistent with our result that the
bias can only be evident for particle numbers $<4\times 10^4$.

>>>>

We then use an Affine Invariant Markov Chain Monte Carlo (MCMC)
implemented in the python module {\texttt{emcee}} \citep{emcee} to sample the
likelihood function distribution.
This runs roughly proportional to $N$. Using a single 2.3Ghz CPU core with two
walkers over 500 steps takes approximatly $0.5$ miliseconds per halo per
particle in the halo, i.e. a halo with $N=2\times 10^{3}$ can be fit
in one second approximately.

From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty. 
We stress that different choices for $\sigma^2_{\rm eff}$ do not
affect the optimal concentration value, only its uncertainty.

@@

Figure \ref{fig:downsampling} summarizes our results.  
The plot on the left shows the average value of $f_{\rm off}$ as a
function of particle number.  
This can be interpreted as the statistical bias on the concentration
estimate. 
For large enough particle numbers, $N_{s}>4\times 10^3$ the results of
the three algorithms show a bias below the $1\%$ level.
For a lower number of particles the results start to deviate.
At $200$ particles the velocity method overestimates the concentration
by a factor $14\%$ while the density method overestimates it by $20\%$.
Around the same sampling scale, the new algorithm shows a more stable
behaviour underestimating the concentration only by a factor of
$1\%$-$3\%$.
Work  that studied the effect of
lowering numerical resolution on the same individual halos
\citep{2008MNRAS.391.1685S} yielded similar values for the
concentration across resolutions.  That work  explored a particle
number range $>5\times 10^5$, completely consistent with our result that the
bias can only be evident for particle numbers $<4\times 10^3$.
